% =====================================================================
%  Appendix
% =====================================================================
%
% Sections in this appendix:
%   A  Upstream patches and reproducibility audit
%   B  Perturbation generation details and example texts
%   C  Probe output invariance check
%   D  Per-backbone bootstrap confidence intervals
%   E  Bit-identical cells: direct verification
%   F  Architectural diagnostic: TaTS's detach() edge
%   G  What about Aurora's pretraining?
%   H  Per-domain breakdown
%   I  Per-horizon breakdown
%
% Tables in this appendix:
%   tab:perbb_*   per-backbone (auto-generated, included from
%                   tab_appendix_perbackbone.tex)
%   tab:perdom_*  per-domain  (auto-generated, included from
%                   tab_appendix_perdomain.tex)
%   tab:perhz_*   per-horizon (auto-generated, included from
%                   tab_appendix_perhorizon.tex)
%   tab:perbackbone   bootstrap CIs, hand-curated
%   tab:perturbation_examples  one example string per condition for one
%                              chosen domain (Health), hand-curated.

\section{Upstream Patches and Reproducibility Audit}
\label{sec:patches}

We pin every upstream repository to a specific commit and apply $13$
small patches via \texttt{code/apply\_repo\_patches.py}.  Patches are
mechanically diffable; this section summarises the categories and their
consequences for reproducibility.

\paragraph{Data-loading correctness.}  \TimeMMD{} ships rows whose
textual columns are sometimes empty strings.  pandas's default
\texttt{read\_csv} converts these to \texttt{NaN}, which several
upstream loaders then silently replace with $0$ or with the string
\texttt{"nan"} -- both of which destroy the meaning of our \Ctwo{}
(\textsc{empty}) condition.  We pass \texttt{keep\_default\_na=False}
in every \texttt{read\_csv} call in the three repos, and we remove a
code path in \MMTSF{} that \texttt{fillna("nan")}s the text column.

\paragraph{API drift.}  All three repos were released against
pandas\,1.x.  We backport four idioms (\texttt{drop(columns=)} instead
of the deprecated positional second argument, \texttt{Series.iloc[]}
casting, etc.) so the same code path runs on pandas\,2.x without
behavioural changes.

\paragraph{Architectural unimodal switch.}  \TaTS{} and \MMTSF{} ship
CLI flags for unimodal evaluation
(\texttt{--text\_emb 0 --prior\_weight 0} and \texttt{--prompt\_weight 0}
respectively).  \Aurora{} hard-codes the text path to always run.  We
add a \texttt{--no\_text} flag to \Aurora{}'s run script that sets
\texttt{text\_input\_ids=None} before calling the model's
\texttt{generate}.  By construction this is the only path the model
exposes for skipping the text branch; \cref{sec:probe_validation}
verifies it matches default-text behaviour bit-exactly on every
\TimeMMD{} cell.

\paragraph{Backbone registration.}  \TaTS{}'s
\texttt{exp/exp\_basic.py} registers only \textsc{iTransformer}.  We
register the eight additional backbones present in the same source
tree (\textsc{Autoformer}, \textsc{DLinear}, \textsc{Informer},
\textsc{FiLM}, \textsc{FEDformer}, \textsc{PatchTST},
\textsc{Crossformer}, \textsc{Transformer}).

\paragraph{Determinism and seeding.}  Each runner sets the model's
RNG via the published seed flag.  We set \texttt{torch.manual\_seed}
and \texttt{numpy.random.seed} per cell.  The orchestrator also fixes
\texttt{CUDA\_VISIBLE\_DEVICES} per shard; we did not pin
\texttt{torch.use\_deterministic\_algorithms(True)} because that
disagrees with one of \MMTSF{}'s flash-attention paths.  \Aurora{}'s
flow-matching head averages $100$ samples for inference; with seeded
sampling we obtain bit-identical predictions per cell across re-runs
(\cref{sec:bit_identical}).

\paragraph{Reproducibility.}  The end-to-end sweep is sharded across
GPUs and resumable; total wall clock for the $7{,}776$-cell sweep is
$\sim 92$ A10G-hours.  The full log of CLI invocations, git SHAs, and
returned MSE/MAE per cell is in our \texttt{sweep\_log.jsonl}
(released alongside the code).

% =====================================================================
\section{Perturbation Generation Details}
\label{sec:perturbations}

\paragraph{Row-level invariants.}  For all data-axis perturbations
(\Ctwo{}--\Cfive{}, \Ceight{}, \Cnine{}) we preserve the row count,
the date column, and the numeric target column.  Our generator
\texttt{code/generate\_perturbations.py} runs a post-hoc validator
that diffs every perturbed CSV against the original on the columns
that should be unchanged.

\paragraph{\Cthree{} (\textsc{shuffled}).}  We permute the text
column within the same domain using a NumPy random generator seeded
by the RunSpec seed.  This is the only condition whose CSV content
depends on the seed; all other conditions live under
\texttt{seed0/} on disk.

\paragraph{\Cfour{} (\textsc{cross-domain}).}  We use a fixed pairing
$f$:\textsc{Agriculture}$\!\to\!$\textsc{Climate},
\textsc{Climate}$\!\to\!$\textsc{Agriculture}, and so on.  For each
target row we pull the text from the source-domain row whose
\texttt{date} is closest to the target row's date (resolving ties
deterministically).  This preserves natural English while breaking
topic alignment.

\paragraph{\Ceight{} (\textsc{oracle}).}  For each row in the test
window we substitute a templated string of the form
\texttt{"The next $H$ values are $y_1, y_2, \ldots, y_H$."}  using
the ground-truth values.  The substitution is applied to train/val
rows -- rolled back by $H$ steps -- so the model sees consistent
oracle structure during fine-tuning (for \TaTS{} and \MMTSF{}).  For
\Aurora{} the test window is what matters.

\paragraph{\Cnine{} (\textsc{zero-priors}).}  We set every entry of
\texttt{prior\_history\_avg} to $0.0$.  \Aurora{} does not consume
this column at all (verified by
\texttt{grep -n prior\_history Aurora/}), so \Cnine{} is by
construction a no-op for \Aurora{} -- exactly the behaviour observed
to bit precision.

\paragraph{Why \Ctwo{} and \Cfive{} also zero the prior.}  Following
the original \TimeMMD{} convention, the perturbations that strip
``text-derived information'' do so for all text-derived numeric
columns as well, including \texttt{prior\_history\_avg}.  This is
what makes \Ctwo{}/\Cfive{}/\Cnine{} a clean three-way comparison:
prior is held at zero in all three, and text varies between empty,
constant, and intact.  If the model reads text at all, these three
should differ.

\paragraph{Examples.}  \Cref{tab:perturbation_examples} shows one
example row from each condition for the \textsc{Public\_Health}
domain (week of 2013-02-04, weighted ILI percentage forecasting).

\input{tab_perturbation_examples}

% =====================================================================
\section{Probe Output Invariance}
\label{sec:probe_validation}

We run \Aurora{}'s \texttt{generate} call twice on a fixed batch with
the same seed: once with no hooks, once with both gradient-norm and
attention-entropy hooks attached.  Forward outputs match
bit-exactly (\texttt{torch.allclose(rtol=0,atol=0)}); the hook
attachment does not perturb the forward graph.  Code in
\texttt{code/aurora\_probes.py::validate\_invariance}.

% =====================================================================
\section{Per-backbone Bootstrap Confidence Intervals}
\label{sec:perbackbone_bootstrap}

A condensed view is given in \cref{tab:perbackbone_compact}, which
reports paired-bootstrap means and $95\%$ confidence intervals against
\Cone{}, broken out per backbone.  All intervals use $10{,}000$
resamples on $108$ paired cells per (model, backbone, condition).
The cluster structure of \cref{tab:main} is preserved across
every backbone: text-content perturbations (\Cthree{}, \Cfour{},
\Ceight{}) have intervals tight to zero, while prior-zeroing
perturbations (\Ctwo{}, \Cfive{}, \Cnine{}) have intervals visibly
above zero.

\begin{table}[h]
\centering
\small
\setlength{\tabcolsep}{2pt}
\renewcommand{\arraystretch}{1.16}
\caption{Mean $\Delta\text{MSE}$ vs.\ \Cone{} with $95\%$ paired
bootstrap CI ($n\!=\!108$ paired cells, $10{,}000$ resamples).
Statistically-significant differences ($p\!<\!0.01$) marked $\star$.
Source: \texttt{summaries/bootstrap\_per\_backbone.csv}.}
\label{tab:perbackbone_compact}
\begin{tabular}{l l r r}
\toprule
\textbf{Model/Bbone} & \textbf{Cond.} & \textbf{$\Delta\overline{\text{MSE}}$} & \textbf{$95\%$ CI} \\
\midrule
\multirow{4}{*}{\Aurora{}}            & \Ctwo{}    & $\phantom{-}0.00202\,\star$ & $[+0.0003,\,+0.0041]$\\
                                      & \Cthree{}  & $\phantom{-}0.00032$        & $[-0.00007,\,+0.0008]$\\
                                      & \Cfour{}   & $-0.00051$                  & $[-0.0028,\,+0.0016]$\\
                                      & \Cnine{}   & $\phantom{-}0$              & $[\phantom{-}0,\,0]$\\
\midrule
\multirow{4}{*}{\MMTSF{}/DLin}        & \Ctwo{}    & $\phantom{-}0.297\,\star$   & $[+0.227,\,+0.374]$\\
                                      & \Cthree{}  & $\phantom{-}0.00071\,\star$ & $[+0.0002,\,+0.0013]$\\
                                      & \Cnine{}   & $\phantom{-}0.298\,\star$   & $[+0.227,\,+0.374]$\\
                                      & \Csix{}    & $\phantom{-}0.331\,\star$   & $[+0.201,\,+0.480]$\\
\midrule
\multirow{4}{*}{\MMTSF{}/Inf}         & \Ctwo{}    & $\phantom{-}0.204\,\star$   & $[+0.120,\,+0.303]$\\
                                      & \Cthree{}  & $\phantom{-}0.018$          & $[+0.001,\,+0.037]$\\
                                      & \Cnine{}   & $\phantom{-}0.211\,\star$   & $[+0.123,\,+0.301]$\\
                                      & \Csix{}    & $\phantom{-}0.201\,\star$   & $[+0.121,\,+0.292]$\\
\bottomrule
\end{tabular}
\end{table}

A full per-(model, backbone, condition) breakdown -- including all
nine conditions and including \MMTSF{}/iTransformer and the three
\TaTS{} backbones -- is provided in \cref{sec:per_horizon}.

% =====================================================================
\section{Bit-Identical Cells: Direct Verification}
\label{sec:bit_identical}

For \Aurora{}, $108/108$ paired cells satisfy $\text{MSE}(\Cone{}) =
\text{MSE}(\Cnine{}) = \text{MSE}(\Csix{})$ to $13$ significant
figures; the bootstrap interval is $[0,0]$ exactly because every
paired difference is exactly zero.  For \TaTS{} on \textsc{DLinear}
and \textsc{iTransformer}, $\Ctwo$, $\Cfive$, and $\Cnine$ give
identical predicted forecasts cell-by-cell because of the
architectural artefact described in \cref{sec:detach_appendix};
these match to $\sim\!13$ significant figures of MSE rather than to
bit precision (small floating-point non-associativity in the
forward pass).

% =====================================================================
\section{Architectural Diagnostic: \TaTS{}'s \texttt{detach()} Edge}
\label{sec:detach_appendix}

In \TaTS{}'s training loop
(\texttt{repos/TaTS/exp/exp\_long\_term\_forecasting.py:519}):
\begin{verbatim}
batch_x = torch.cat(
    [batch_x, prompt_emb], dim=-1
).detach()
\end{verbatim}
\noindent
The \texttt{.detach()} severs gradient flow into \texttt{prompt\_emb},
which is the output of \texttt{self.mlp(text\_embeddings)}.
\TaTS{}'s training loop nevertheless instantiates a separate optimiser
for \texttt{self.mlp} and calls \texttt{model\_optim\_mlp.step()}
every iteration -- but those parameters never receive gradients.

\paragraph{Consequence.}  The text-projection MLP $\psi$ stays at
random initialisation throughout fine-tuning.  The downstream backbone
receives a random projection of frozen-LLM features as additional
input channels; over training it learns to assign small or zero
weight to those channels because they cannot reduce loss reliably.
At test time, with text content varied (\Ctwo{}/\Cfive{}/\Cnine{})
the random projection maps to slightly different vectors but the
trained backbone has already collapsed those channels to near-zero
contribution.  This is a sufficient mechanical explanation for the
bit-identical behaviour we observe in \cref{tab:main}.

\paragraph{Significance.}  This is not a takedown of \TaTS{} as a
research artefact; the released code may differ from the authors'
internal codebase.  We highlight it because anyone re-deriving a
``multimodal'' lift from \TaTS{}'s public code is, structurally,
re-deriving the prior-residual gap rather than a text gap.

% =====================================================================
\section{What about Aurora's Pretraining?}
\label{sec:aurora_pretrain}

\Aurora{}'s zero-shot regime means our findings concern the released
\emph{checkpoint}'s use of text on \TimeMMD{}.  We do not re-pretrain
\Aurora{}; we cannot rule out training-time contamination of
\TimeMMD{} content into the pretraining corpus.  The fact that
$\Cone{}/\Cnine{}/\Csix{}$ are bit-identical and $\Ctwo{}/\Cfive{}$
are within $0.03\%$ of $\Cone{}$ tells us that, conditional on the
released checkpoint, text content is not a behavioural lever on this
benchmark.  Our claim is conditional on these checkpoints; we do not
claim that no text--time-series foundation model can use text.

% =====================================================================
\section{Per-backbone, Per-domain, Per-horizon Tables}
\label{sec:per_horizon}

The next three subsections expand \cref{tab:main} along three
axes: (i) per backbone, (ii) per domain, and (iii) per horizon.  In
every breakdown, the two-cluster pattern from the main paper is
preserved.  Tables are auto-generated from the locked summaries by
\texttt{paper/figures/make\_appendix\_tables.py}.

\subsection{Per-backbone}
\label{sec:perbackbone_full}

\input{tab_appendix_perbackbone}

\subsection{Per-domain}
\label{sec:perdomain_full}

\input{tab_appendix_perdomain}

\subsection{Per-horizon}
\label{sec:perhorizon_full}

The horizons available depend on the source domain's sampling
frequency: \textsc{Environment} (daily) uses
$\{48, 96, 192, 336\}$, \textsc{Health}/\textsc{Energy} (weekly) use
$\{12, 24, 36, 48\}$, and the remaining six monthly domains use
$\{6, 8, 10, 12\}$.  The table below merges across domains, so
columns with a particular horizon $H$ aggregate over only the
domains that include that horizon.

\input{tab_appendix_perhorizon}
